\section{Method}
\label{sec:method}

As the trajectories of the \cite{chen2025robo2vlm} dataset are success only, this framework is designed to use temporal and action-phase-label - vision alignment to test model capabilities across object-, state-, manipulation-, and scene-understanding. And in particular temporal comprehension of robotic tasks. 

\subsection{Metrics}
We measure consistency as whether
independently elicited answers agree across views, using three metrics: TL$\leftrightarrow$TR
agreement, agreement of at least one top view with the wrist view
($\geq$1~Top$\leftrightarrow$W), and All-3-agree, where all three views give the same
answer. Treating each view as an independent binary draw gives baselines of 50\%, 75\%,
and 25\% for the three metrics, respectively.

\subsection{Models}
We evaluate seven publicly available VLMs spanning three model families, chosen to cover
a range of parameter counts within and across families: Gemma3~\cite{gemma2025gemma3}
(E2B, E4B, 31B), Qwen3~\cite{qwen2025qwen3} (4B, 8B, 32B), and
InternVL3~\cite{chen2025internvl3} (14B). All models are run zero-shot on the ETH Zurich
student cluster without any task-specific training.

\subsection{Zero-Shot Evaluation Setup}
The models are prompted with a single image (or two, for \texttt{progress}), the goal
text, and a multiple-choice question with answer options labeled A through E (the exact
count depends on the sub-task). We extract the first letter predicted in a single forward
pass as the model's answer, without few-shot examples or task-specific instructions. In
the chain-of-thought (CoT) variant, the model is instead asked to reason freely before
stating a final answer letter; we apply the same extraction procedure to the last letter
emitted in the response. \TODO{include the exact prompt template here, or move it to
supplementary.}

\subsection{LoRA Fine-Tuning}
We additionally fine-tune Qwen3-4B with LoRA to control for the possibility that low
zero-shot accuracy reflects missing task-specific grounding rather than a fundamental
reasoning ceiling. We use rank 16, alpha 32, and dropout 0.05 on the query, key, value,
and output projections of the language-model layers, trained on a scene-level split with
20 training scenes (2535 questions) and 7 validation scenes (983 questions); the held-out
test split (20 scenes, 2706 questions) shares no scene with training, preventing the
model from memorizing specific visual scenes. Training runs for 3 epochs with a learning
rate of $2\times10^{-4}$, batch size 1 with gradient accumulation 8 (effective batch size
8), and a maximum sequence length of 1024 tokens (seed 42). We compare against zero-shot
Qwen3-4B evaluated on the identical 2706-question test split, giving an apples-to-apples
BASE vs.\ +LoRA comparison.
