\section{Discussion}
\label{sec:discussion}

\textbf{Accuracy Analysis.} Object state --- such as contact or orientation change --- and temporal understanding are underrepresented in web-scale VLM training data relative to object identification and position estimation. Models can usually describe which objects are present and where, but not reliably what has happened to them or which phase of an action is depicted. The vision backbone encoding is not influenced by the prompt; the encoded signal must therefore suffice for answering on its own.
High \texttt{task\_success} accuracy in the Qwen3 family (70\%+) cannot be safely attributed to genuine perception, as these models may exhibit high-confidence convergence on the majority label.

\textbf{Scale difference.} Scaling consistently improves
accuracy across all families, but even 31B/32B-class models plateau in the
high-40s\% on phase-level reasoning. Cross-family comparison shows, this hypothesis holds, Qwen3-8B is competitive with InternVL3-14B despite the parameter gap, suggesting architecture and training data matter at least as much as raw parameter count. From these results it is difficult to estimate, whether vision backbone model, data training, or vision-language alignment are the main bottleneck. Nonetheless, it is unlikely that more parameters alone close the gap, but rather robot-domain training data is needed.

\textbf{LoRA fine-tuning.} Although LoRA improves overall accuracy by 11pp, the decomposition in \cref{sec:results-lora} shows that roughly 80\% of the gain is attributed to two coarse heuristics rather than stronger perception capabilities. The model learns to detect if a random scene compared to the task is shown, implying that the vision-backbone signal can be leveraged effectively for such detection. This presents a generalizable skill rather than memorization, when compared between training, validation or test on subtasks action\_phase\_id and task\_success. Yet it is domain-mismatch detection, not phase or temporal reasoning. The other gain is in \texttt{phase\_success}, specifically at MIDDLE phase positions, but as this occurs in questions where the ordered list of action phases is given, we believe the model is improving at guessing from the given list context rather than from the image itself.
%TODO control if gain in action phase detection is stronger in action phase context giving
% TODO check phase success gt label distribution
This \texttt{phase\_success} gain is net positive under this dataset's label distribution, but a regression for the most safety-critical use case: detecting when a robot is currently in the middle of an action. The results show that at this scale no improvement of reasoning can be seen, but mostly better dissecting of the vision-backbone signal.

\textbf{CoT prompting.}
Based on the traces the model exhibits capabilities to describe and reason over the different views, but it struggles with the basics, such as aligning the views correctly, for correct depth estimation for examples.
Single-letter prompts force a direct guess that happens to land closer to the
skewed label distribution (72\% ``No'' on \texttt{task\_success}), while CoT makes the
model's reasoning more careful and favors expressing uncertainty. Under the assumption that the model's capabilities can be judged from its reasoning, Gemma3-E2B and E4B show a lack of clear internal representation of the given scenes.

\textbf{Multi-view consistency tracks confidence, not correctness.} Models are most self-consistent when uniformly right (72.3\%) or uniformly wrong (66.4\%) across views, and least consistent when only some views are correct (34.4\%). We read this as
view-consistency reflecting how unambiguous a scene is to the model, not whether its
answer is correct. This connects to the scene-difficulty finding (\cref{sec:results}):
scenes that drive accuracy down also destabilize cross-view agreement, suggesting both
effects share the same root causes. For multi-view fusion, this implies that simple voting across views may not help on exactly the hardest scenes, where inconsistency is highest.
