\section{Conclusion}
\label{sec:conclusion}

We benchmark VLMs on two robot-supervision capabilities and arrive at three findings.
First, zero-shot state reasoning is weak: accuracy stays only marginally above random across model families, scale helps within a family, but even 30B+ models
plateau well short of reliable supervision. Second, chain-of-thought prompting introduces additional difficulties and requires significantly more compute; free-form reasoning chains appear to introduce hallucinated conclusions on perceptual tasks rather than improving them. Third, fine-tuning gains are real but mostly coarse: LoRA lifts Qwen3-4B accuracy by +10.9pp, but the gains cannot be attributed to improved temporal or state reasoning. Our
results suggest current VLMs are not yet ready for reliable robot-state supervision out of the box, and that fine-tuning gains must be carefully decomposed by question type and answer class before they can be trusted for deployment.

\section{Limitations \& Future Work}
\label{sec:limitations}

\textbf{Limitations.} Manual annotation limits scale: 47 scenes is enough for a
controlled study but too small to draw strong statistical conclusions about
model-family differences; an automated annotation pipeline as in \cite{chen2025robo2vlm} would let this scale up. Train and test splits are drawn from the DROID dataset, so our results do not test sim-to-real or cross-dataset generalization. Finally, LoRA fine-tuning was only run on Qwen3-4B, so it is unknown whether the CnBD-generalization and IPI-collapse pattern holds
at larger scale.

\textbf{Future work.} We plan to use the Robo2VLM-1 pipeline to generate a larger-scale dataset for more robust cross-difficulty testing. Running LoRA on larger models and fine-tuning the vision backbone would clarify whether the gain-decomposition pattern holds at scale. Running CoT on larger models would separately test whether the over-caution effect is size-dependent. Finally, another interesting angle would be to use video segments around keypoints of the model, by either using video-language-models directly or by giving multiple images in close temporal consistency. This could surface new temporal reasoning capabilities that our single-frame queries can't test.

\section{Statement of work distribution.}
While we all contributed to each stage of the project, we each spent more time on certain tasks:
Christian dedicated the most time to LoRA training, the data labeling interface, and compiling the labeled data into the final benchmark questions.
Alexander contributed most of his time to testing models on the GB10s, finalizing the report, and presenting our intermediate results during the semester.
Ana contributed most of her time to understanding the DROID dataset and pipeline, preparing the mid-term presentations, and writing the poster with graphics.
Defne contributed most of her time to finding flaws in the original DROID dataset, benchmarking the models, and understanding their failure modes.
We all equally contributed to labeling the data.